---
title: "Naive Bayes"
output: html_document
---

```{r}
library(mlbench)
data(HouseVotes84)  
```

barplots for specific issue

```{r}
plot(as.factor(HouseVotes84[,2]))
title(main='Votes cast for issue 1', xlab="vote", ylab= "Num reps" )
```

By party
```{r}

Repub <-  HouseVotes84$Class=="republican"
Democrat <- HouseVotes84$Class=="democrat"
```

```{r}
plot(as.factor(HouseVotes84[Repub,2]))
title(main="Republican votes cast for issue 1", xlab="vote", ylab="Num reps")
plot(as.factor(HouseVotes84[Democrat,2]))
title(main="Democrat votes cast for issue 1", xlab="vote", ylab="Num reps")

```

We have a lot of NA. We are going to impute the values

Functions needed for imputation

function to return number of NAs by vote and class (democrat or republican)
```{r}
na_by_col_class <- function (col,cls){return(sum(is.na(HouseVotes84[,col]) & HouseVotes84$Class==cls))}

```

function to compute the conditional probability that a member of a party will cast a "yes" vote for a particular issue. The probability is based on all members of the party who actually cast a vote on the issue (ignores NAs).

```{r}
p_y_col_class <- function(col,cls){
  sum_y<-sum(HouseVotes84[,col]=="y" & HouseVotes84$Class==cls,na.rm = TRUE)
  sum_n<-sum(HouseVotes84[,col]=="n" & HouseVotes84$Class==cls,na.rm = TRUE)
  return(sum_y/(sum_y+sum_n))
}

```

Check the prob of yes vote by a democrat in issue 5
```{r}
p_y_col_class(5,"democrat")
```

impute missing values.
 If the republican congresman didn't vote, then we are allocating 'y' or 'n' based on if their
 party voted 'y' or 'n'

```{r}
for (i in 2:ncol(HouseVotes84)) {
  if(sum(is.na(HouseVotes84[,i])>0)) {
    c1 <- which(is.na(HouseVotes84[,i])& HouseVotes84$Class=="democrat",arr.ind = TRUE)
    c2 <- which(is.na(HouseVotes84[,i])& HouseVotes84$Class=="republican",arr.ind = TRUE)
    HouseVotes84[c1,i] <-
      ifelse(runif(na_by_col_class(i,"democrat"))<p_y_col_class(i,"democrat"),"y","n")
    HouseVotes84[c2,i] <-
      ifelse(runif(na_by_col_class(i,"republican"))<p_y_col_class(i,"republican"),"y","n")}
}
```

divide into test and training sets
create new col "train" and assign 1 or 0 in 80/20 proportion via random uniform dist


```{r}
HouseVotes84[,"train"] <- ifelse(runif(nrow(HouseVotes84))<0.80,1,0)

```

get col number of train / test indicator column (needed later)

```{r}
trainColNum <- grep("train",names(HouseVotes84))
```

separate training and test sets and remove training column before modeling
```{r}
trainHouseVotes84 <- HouseVotes84[HouseVotes84$train==1,-trainColNum]
testHouseVotes84 <- HouseVotes84[HouseVotes84$train==0,-trainColNum]

```

load e1071 library and invoke naiveBayes method
```{r}
library(e1071)
nb_model <- naiveBayes(Class~.,data = trainHouseVotes84)

nb_model

```

Lets test the model
```{r}

nb_test_predict <- predict(nb_model,testHouseVotes84[,-1])

```

fraction of correct predictions
```{r}
mean(nb_test_predict==testHouseVotes84$Class)

#confusion matrix
table(pred=nb_test_predict,true=testHouseVotes84$Class)
```

