---
title: "Random Forest"
author: "Amruta"
date: "8/6/2019"
output: html_document
---

* Clear Environment variables
```{r}
rm(list = ls(all = T))
```

* Load the required libraries
```{r}
library(DMwR)
library(randomForest)
library(caret)
```

* Read the data into R * since thr is no header given initially in the file, we give header 
```{r}
data = read.table('hepatitis.txt', header=F, dec='.',
                  col.names=c('target','age','gender','steroid',
                              'antivirals','fatigue','malaise',
                              'anorexia','liverBig','liverFirm',
                              'spleen','spiders','ascites',
                              'varices','bili','alk','sgot',
                              'albu','protime','histology'), 
                  na.strings=c('?'), sep=',')

```

* Understand the data 
```{r}
str(data)
summary(data)
```

* check target count  1: Die; 2: Live
```{r}
table(data$target)
str(data$target)  
```

* Convert 1s and 2s into 1s and 0s
```{r}
data$target= ifelse(data$target==1, 1, 0 )  
```

* The numerical variables are: age, bili, alk, sgot, albu and protime
  The categorical variables are: the remaining 14 variables
```{r}

num_Attr = c("age", "bili", "alk", "sgot", "albu", "protime")
cat_Attr = setdiff(names(data), num_Attr)

```

* Seperate numerical and categorical variables and convert them into appropriate type
```{r}
cat_Data = data.frame(sapply(data[,cat_Attr], as.factor))
num_Data = data.frame(sapply(data[,num_Attr], as.numeric))
data = cbind(num_Data, cat_Data)
rm(num_Attr, cat_Attr)
rm(cat_Data, num_Data)

```

* Split dataset into train and test
```{r}
set.seed(9)

train_RowIDs = sample(1:nrow(data), nrow(data)*0.7)
train_Data = data[train_RowIDs,]
test_Data = data[-train_RowIDs,]
rm(train_RowIDs)

```

* Check how records are split with respect to target attribute.
```{r}
table(data$target)
table(train_Data$target)
table(test_Data$target)
rm(data)
```

  As part of Pre-processing, Imputation and scaling are done after train-evaluation/test split.Check to see if missing values in data
```{r}
sum(is.na(train_Data))
sum(is.na(test_Data))

```

* Imputing missing values using KNN
```{r}
train_Data <- knnImputation(data = train_Data, k = 5)
sum(is.na(train_Data))
test_Data <- knnImputation(data = test_Data, k = 5, distData = train_Data)
sum(is.na(test_Data))

```

* Model Building

  Build the classification model using randomForest
```{r}
set.seed(123)

model = randomForest(target ~ ., data=train_Data, 
                     keep.forest=TRUE, ntree=100) 

```

* Print and understand the model
```{r}
print(model)
```

*  Important attributes
```{r}
model$importance  
round(importance(model), 2)  
```

* Extract and store important variables obtained from the random forest model
```{r}

rf_Imp_Attr = data.frame(model$importance)
rf_Imp_Attr = data.frame(row.names(rf_Imp_Attr),rf_Imp_Attr[,1])
colnames(rf_Imp_Attr) = c('Attributes', 'Importance')
rf_Imp_Attr = rf_Imp_Attr[order(rf_Imp_Attr$Importance, decreasing = TRUE),]

```


*  plot (directly prints the important attributes) 
```{r}
varImpPlot(model)
```

*  Predict on Train data '
```{r}
pred_Train = predict(model, 
                     train_Data[,setdiff(names(train_Data), "target")],
                     type="vote", 
                     norm.votes=TRUE) 

pred_Train = predict(model, 
                     train_Data[,setdiff(names(train_Data), "target")],
                     type="response", 
                     norm.votes=TRUE) #normalize the response
```

*  Build confusion matrix and find accuracy   
```{r}
cm_Train = table("actual"= train_Data$target, "predicted" = pred_Train);
accu_Train= sum(diag(cm_Train))/sum(cm_Train)
rm(pred_Train, cm_Train)
```

*  Predicton Test Data
```{r}
pred_Test = predict(model, test_Data[,setdiff(names(test_Data),
                                              "target")],
                    type="response", 
                    norm.votes=TRUE)
```

*  Build confusion matrix and find accuracy  

```{r}
cm_Test = table("actual"=test_Data$target, "predicted"=pred_Test);
accu_Test= sum(diag(cm_Test))/sum(cm_Test)
rm(pred_Test, cm_Test)

accu_Train 
accu_Test
```

*  Build randorm forest using top 11 important attributes. 
```{r}
top_Imp_Attr = as.character(rf_Imp_Attr$Attributes[1:12])

set.seed(15)
```

*  Build the classification model using randomForest

```{r}
model_Imp = randomForest(target~.,
                         data=train_Data[,c(top_Imp_Attr,"target")], 
                         keep.forest=TRUE,ntree=100) 
```

*  Print and understand the model
```{r}
print(model_Imp)
```

*  Important attributes
```{r}
model_Imp$importance  
```

*  Predict on Train data 

```{r}
pred_Train = predict(model_Imp, train_Data[,top_Imp_Attr],
                     type="response", norm.votes=TRUE)
```

*  Build confusion matrix and find accuracy   
```{r}
cm_Train = table("actual" = train_Data$target, 
                 "predicted" = pred_Train);
accu_Train_Imp = sum(diag(cm_Train))/sum(cm_Train)
rm(pred_Train, cm_Train)
```

*  Predicton Test Data
```{r}
pred_Test = predict(model_Imp, test_Data[,top_Imp_Attr],
                    type="response", norm.votes=TRUE)
```

*  Build confusion matrix and find accuracy   
```{r}
cm_Test = table("actual" = test_Data$target, 
                "predicted" = pred_Test);
accu_Test_Imp = sum(diag(cm_Test))/sum(cm_Test)
rm(pred_Test, cm_Test)

accu_Train
accu_Test
accu_Train_Imp
accu_Test_Imp
```

* Select mtry value with minimum out of bag(OOB) error. selection of number of features.
```{r}
mtry <- tuneRF(train_Data[-7],train_Data$target, ntreeTry=100,
               stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
```

* how many step u need to jump while selecting features * if u dont see impruvement by given value stop.

```{r}
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
print(mtry)
print(best.m)
```

* Build Model with best mtry again - 

```{r}
set.seed(71)
rf <- randomForest(target~.,data=train_Data, mtry=best.m, importance=TRUE,ntree=100)
print(rf)
```

* Evaluate variable importance
```{r}
importance(rf)
```

*  Important attributes
```{r}
model$importance  
round(importance(model), 2)   
```

*  Extract and store important variables obtained from the random forest model
```{r}
rf_Imp_Attr = data.frame(model$importance)
rf_Imp_Attr = data.frame(row.names(rf_Imp_Attr),rf_Imp_Attr[,1])
colnames(rf_Imp_Attr) = c('Attributes', 'Importance')
rf_Imp_Attr = rf_Imp_Attr[order(rf_Imp_Attr$Importance, decreasing = TRUE),]
```

*  Predict on Train data 
```{r}
pred_Train = predict(model, 
                     train_Data[,setdiff(names(train_Data), "target")],
                     type="response", 
                     norm.votes=TRUE)
```

*  Build confusion matrix and find accuracy   
```{r}
cm_Train = table("actual"= train_Data$target, "predicted" = pred_Train);
accu_Train = sum(diag(cm_Train))/sum(cm_Train)
rm(pred_Train, cm_Train)
```


*  Predicton Test Data
```{r}
pred_Test = predict(model, test_Data[,setdiff(names(test_Data),
                                              "target")],
                    type="response", 
                    norm.votes=TRUE)
```

*  Build confusion matrix and find accuracy   
```{r}
cm_Test = table("actual"=test_Data$target, "predicted"=pred_Test);
accu_Test= sum(diag(cm_Test))/sum(cm_Test)
rm(cm_Test)

accu_Train
accu_Test
```

Note: You can find data at : "http://archive.ics.uci.edu/ml/datasets/Hepatitis